---
title: "Screening24"
output: pdf_document
date: "2025-05-08"
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
*Functions* 
The following function will replace all "<LOD"
- har det no ÃƒÂ¥ si om det er mellomrm mellom < og tallet? nei, tror ikke det
```{r}
#function to recognize "<LOD" and to replace with 0.5*LOD. The next step is to use this only for combinations of sample type and site with LOD >30
dLimity <- function(v){
  v <- as.character(v)
  isLimit <- grepl("<", v)  #grepl() is a built-in R function looks for a smaller string of characters
  v[isLimit] <- as.numeric(gsub("<(.*)", "\\1", v[isLimit]))/2 #gsub(pattern, replacement, x) for substitution
  as.numeric(v)
}

#Test to set desired number of digits
sigfig <- function(vec, digits){
  return(gsub("\\.$", "", formatC(signif(vec,digits=digits), digits=digits, format="fg", flag="#")))
}
```

```{r, results=FALSE, message=FALSE, warning=FALSE}
#install.packages(c("dplyr", "ggplot2", "plyr", "Rmisc", "tidyverse", "lubridate", "stringr"))
Packages <- c("plyr", "dplyr", "ggplot2",  "tidyr", "egg", "reshape2", "forcats", "RColorBrewer")
lapply(Packages, library, character.only = TRUE)
#trengs disse? "Rmisc", "tidyverse", "lubridate", "stringr",
```

#Load file
```{r echo = T, results = 'hide'}
Sys.setlocale(locale="no_NO")
df24 <- read.table("Scr24_Data_05_V03.txt", skip=0, header=TRUE, sep="\t", na.string=c(""))
```

#Make summary table count - number of samples per sample type and compound
```{r}
count <- df24 %>% count(KOMPONENT, Sample.type)
head(df24)

#add grouping
df24x = df24%>%
  mutate(Level = case_when(KOMPONENT %in% c("Cadmium", "Lead", "Arsenic", "Nickel", "Cobalt", "Iron", "Manganese", "Chromium", "Vanadium", "Potassium", "Aluminium", "Magnesium", "Silver") ~ "Other metals",
                           KOMPONENT %in% c("Diflubenzuron", "Teflubenzuron", "Emamectin") ~ "Lice remediation",
                                KOMPONENT %in% c("TFA", "CFPCN", "ACBE", "CHPAPE", "ACCA", "CDHFPCN", "CHFPCN", "CCEFPE", "CCPCA", "BCCPCA") ~ "Degradation products - Tralopyril",
                           KOMPONENT %in% c("Tralopyril", "Copper pyrithione", "Zinc pyrithione", "Copper", "Zinc") ~ "Antifouling agents"))



df24x <- subset(df24x, KOMPONENT != "Grunnpris")
```


```{r Detection frequency}
#remove n.a due to limited sample material. SHould not be included in calculations
#df24y <- subset(df24x, RESULTAT != "n.a.")

df24y_freq = df24x %>%
  group_by(KOMPONENT, Sample.type) %>%
  mutate(lodcount = sum(lengths(regmatches(RESULTAT, gregexpr("<", RESULTAT)))), 
            n = (length(RESULTAT)),
            det_frac = (100-(lodcount/n)*100))%>%
  ungroup() 

```


```{r Avoid plotting average of those with only < LOD}

df24z_freq <- df24y_freq %>% 
  mutate(plotting = ifelse(det_frac == 0, 
                           "<LOD", 
                           RESULTAT))

# claculate 0.5xLOD 
df24z_filt_X <- cbind(df24z_freq[,c(1:27)], apply(df24z_freq[,28, drop=F],2, dLimity))

```

Calculate mean
- sjekk at ble riktig! ta ut excel f eks
```{r}
df24z_finalx = df24z_filt_X  %>%
  group_by(KOMPONENT, Sample.type) %>%
  mutate(meanx = mean(plotting, na.rm = TRUE))%>% 
  ungroup()

```

##geom_text(aes(label=ifelse(is.na(meanx), "", sigfig(meanx, 2))), color = "white", size = 5)+
#geom_text(aes(label=ifelse(is.na(meanx), "", round(meanx, digits=3))), color = "white", size = 5)

Set order of substances in plot
```{r}
df24z_finalx$KOMPONENT <- as.character(df24z_finalx$KOMPONENT)
df24z_finalx$Level <- as.character(df24z_finalx$Level)


```

##Part 1
```{r}
#to get "s" for suspect and "n.a." for not available
df24z_finalxy <- df24z_finalx %>% 
  mutate(meanxyz = ifelse(KOMPONENT == "HCPPE" |
                            KOMPONENT == "ACCA" | 
                            KOMPONENT =="CCEFPE"
                          | KOMPONENT =="CDHFPCN"|
                            KOMPONENT == "CFPCN"|
                            KOMPONENT == "CHFPCN"|
                            KOMPONENT =="CHPAPE",
                           "S", 
                           " "))


#Rename names e.g. starting with a number
#unique(part1_finalxy$compound)

#part1_finalxyx <- subset(part1_finalxy, compound!="DTBPPO")
#part1_finalxy$compound <- revalue(part1_finalxyx$compound, c("BENAZOL.P"="BENAZOL P", #"X23BPDT"="23BPDT", "X34DMPZ"="34DMPZ", "X2BBTMP"="2BBTMP", "X34DMPZP"="34DMPZP", #"X35DMPZ"="35DMPZ", "X2BDOMP"="2BDOMP", "X4NAS"="4NAS", "B.2ETF"="B-2ETF", "X2BHMPTP"="2BHMPTP", #"X2BMPP"="2BMPP", "X2BMTP"="2BMTP", "X22BPDT"="22BPDT", "X2TBCLBEP"="2TBCLBEP", "B.2ETF"="B-2ETF", #"X2BDMP"="2BDMP", "DTBPPO.1"="DTBPPO" )) #er den siste riktig?

#unique(part1_finalxy$compound)
#part1_finalxyz <- subset(part1_finalxy, compound!="Fett")

#length(unique(part1_finalxy$compound))
```

### All together
```{r }
#subset
#wwt <- subset(part1_finalxyz, Category == "Wastewater treatment")

#better names for sample types
#unique(wwt$Sample.type)

#wwt$Sample.type <- revalue(wwt$Sample.type, c("Water, filtered" = "Water, filtered (ng/L)", "Water, particulate" = "Water, particulate (ng/g, d.w.)", "Sludge" ="Sludge (ng/g, d.w.)", "Sediment" ="Marine sediment (ng/g, d.w.)", "Blue mussel" ="Blue mussel (ng/g, w.w.)"))

#set order of dample types
#set order of dample kcompounds
df24z_finalxy <- df24z_finalxy %>% 
  mutate(Sample.type = factor(Sample.type, levels = c("Seawater", "Sediment", "Benthic animals", "Fish gills", "Fish liver")))

#set order of dample kcompounds
df24z_finalxy <- df24z_finalxy %>% 
  mutate(KOMPONENT = factor(KOMPONENT, levels = c("Copper pyrithione", "Copper", "Zinc pyrithione", "Zinc", "Tralopyril",
    "ACBE", "ACCA", "BCCPCA", "CCEFPE", "CCPCA", "CDHFPCN", "CFPCN", "CHFPCN", "CHPAPE", "TFA", 
    "Emamectin", "Diflubenzuron", "Teflubenzuron",
    "Aluminium", "Arsenic", "Cadmium", "Chromium", "Cobalt", "Iron", "Lead", "Magnesium", "Manganese", "Nickel", "Potassium", "Silver", "Vanadium")))


```

 
```{r }
#Fjern alle < LOD
#wwt3 <- subset(wwt2, meanx != "NaN") 
df24z_finalxyz <- subset(df24z_finalxy, det_frac != 0) 
df24z_finalx

ggplot(df24z_finalxy, aes(x=Sample.type, y=fct_rev(KOMPONENT), fill=det_frac)) +
  geom_tile(aes(fill = det_frac), colour="black")+  # background colours are mapped according to the value column
  #geom_text(aes(label=ifelse(is.na(meanx), "", sigfig(meanx, 1))), color = "white", size = 5)
  geom_text(aes(label= ifelse(meanxyz=="S", meanxyz, ifelse(is.na(meanx), " ", as.numeric(sigfig(meanx, 1))))), color = "white", size = 5)+
 scale_fill_gradient2(low = "#74C476", 
                       mid = "orange", 
                       high = "red",
                       midpoint = 50,
                       breaks=c(0,50,100), limits=c(0.5,100), na.value = "lightgrey")+  # determine the colour
   theme(panel.grid.major.x=element_blank(), #no gridlines
        panel.grid.minor.x=element_blank(), 
        panel.grid.major.y=element_blank(), 
        panel.grid.minor.y=element_blank(),
        panel.background=element_rect(fill="white"), # background=white
        axis.text.x = element_text(angle=45, hjust = 1,vjust=1,size = 15,colour="black"),
        plot.title = element_text(size=18,colour="black", vjust = 1, face="bold"),
        axis.text.y = element_text(size = 12, colour="black"),
        legend.text=element_text(size=13),
        axis.line.x = element_line(color="black", linewidth = 1),
        axis.line.y = element_line(color="black", linewidth = 1),
        axis.ticks = element_line(color="black", linewidth=1),
        legend.position =c(0.85, 0.96),
        legend.direction = "horizontal",
        plot.margin = margin(1.3, 0.2, 0.15, 0.15, "cm"), #top, right, bottom, left
        legend.margin=margin(0,0,0,0),
        legend.box.margin=margin(10,10,140,10),
        strip.text.x = element_text(size = 15),
        panel.spacing.x = unit(0.2, "lines"),
        legend.title=element_text(size=12), 
        strip.text.y = element_text(angle = 0, size=12)) +
  ggtitle("") + 
  scale_x_discrete(name="") +
  scale_y_discrete(name="") +
  labs(fill="Detection frequency (%)")+
  facet_grid(Level~., scales="free", space="free_y", labeller = label_wrap_gen(width=2))


ggsave("WW_t2.png", wwt_long2, width = 10, height = 12)

```

##Antifouling agents
- distance from site
- sum of degradation products and traloporyl
```{r}
#Tralopyril distance from fishfarm and different matrics
tr <- subset(df24z_finalxy, KOMPONENT == "Tralopyril" & Sample.type %in% c("Seawater", "Sediment", "Benthic animals"))

tr2 <- subset(df24z_finalxy, Level %in% c("Antifouling agents", "Degradation products - Tralopyril") & Sample.type %in% c("Seawater", "Sediment", "Benthic animals"))

tr2x <- subset(tr2, !KOMPONENT %in% c("Copper pyrithione", "Zinc pyrithione", "Copper", "Zinc"))

#add grouping
tr2y = tr2x %>%
  mutate(Pot = case_when(Level %in% c("Antifouling agents") ~ "Original",
                           Level %in% c("Degradation products - Tralopyril") ~ "Degradation"))

    
#fill = interaction(Subcategory, Group))

library(dplyr)
library(stringr)
tr2z=tr2y %>% 
  mutate(x_label = factor(str_replace(interaction(Pot, KOMPONENT), '\\.', ' / '),
                          ordered=TRUE))

ggplot(df, aes(x=x_label, y=total, fill=type)) +
  geom_bar(stat='identity') +
  labs(x='Year / Treatment')


tr2zz = tr2z  %>%
  group_by(KOMPONENT, Sample.type, Plass) %>%
  mutate(meanx2 = mean(plotting, na.rm = TRUE))%>% 
  ungroup()

#set order of dample kcompounds
tr2zzz <- tr2zz %>% 
  mutate(Pot = factor(Pot, levels = c("Original", "Degradation")))


ggplot(tr2zzz, aes(x=as.factor(Pot), y=as.numeric(meanx2)+1, fill= forcats::fct_rev(KOMPONENT)))+
  #geom_point()+
  geom_bar(stat="identity", position="stack")+
  #geom_col( width=0.8, position=position_dodge())+
  #geom_bar(stat="identity", position = position_dodge())+
  scale_y_continuous(trans='log10')+
  #geom_dotplot(binaxis='y', stackdir='center')
  facet_grid(Sample.type~Plass, scales="free")

#in Tralopyril, look for effects from station. Is there higher in Mørenot?
ggplot(tr, aes(x=as.factor(Plass), y=as.numeric(plotting), fill=Lokasjon, colour=Lokasjon))+
  geom_point(size=3)+
  scale_y_continuous(trans='log10')+
  #geom_dotplot(binaxis='y', stackdir='center')
  facet_grid(Sample.type~., scales="free")
  ylim(0, 100)
  
#tralopyril with depth - only in sediment
t_sed <- subset(tr, Sample.type =="Sediment")  

ggplot(t_sed, aes(x=as.numeric(Havdyp), y=as.numeric(plotting), fill=Lokasjon, colour=Lokasjon))+
  geom_point(size=3)+
  scale_y_continuous(trans='log10')
  #geom_dotplot(binaxis='y', stackdir='center')
  facet_grid(Sample.type~., scales="free")
  ylim(0, 100)

```


#Looking at Cu and Zn

